---
title: "Assign WoS categories to journals"
author: 
- Damiano Oldoni
- Rien De Raedt
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Setup
 
## Load libraries
 
```{r load_libraries, message = FALSE}
library(tidyverse)  # to do data science
library(here)       # to find files
```

## File paths

Specify the directory containing text files:

```{r path_file}
files_dir <- here("data", "raw")
```

Each category is a text file. All text files start with `SCIENCE CITATION INDEX EXPANDED - `

```{r extract_names_from_dir}
pattern <- "SCIENCE CITATION INDEX EXPANDED - "
files_all <- dir(files_dir)
files <- files_all[which(str_detect(files_all, pattern = pattern))]
# examples
files[1:10]
```

# Load and tidy data

## Import journals in categories

Import the text files containing the journals for each category as a list of raw dataframes:

```{r import_data, warning=FALSE}
journals_raw <- map(files, ~ read_tsv(
  str_c(files_dir, ., sep = "/"),
  col_names = FALSE, 
  col_types = cols(X1 = col_character())))
```

We extract the category from very first line of each file and assign it as names of each element of the list:

```{r extract_categories}
categories <- map(
  journals_raw, function(df) {
    unlist(str_split(df$X1[1], pattern = " - "), 
           recursive = TRUE)[2]
  }
)
names(journals_raw) <- categories
```

We extract the journals by selecting rows starting with a number followed by a dot and a space, e.g. `1. ACTAS UROLOGICAS ESPANOLAS`:

```{r extract_journals}
journals_category <- map(
  journals_raw, function(df) {
    df %>%
      filter(str_detect(X1, "^[0-9]*\\. "))
   }
)
```

We merge the list of dataframes, we add category as second column and we remove the pattern number + dot + space from each row. For example,  `1. ACTAS UROLOGICAS ESPANOLAS` becomes `ACTAS UROLOGICAS ESPANOLAS`:

```{r add_column_category_merge_df}
journals_df <- map2_dfr(
  journals_category, names(journals_category), function(df, cat) {
    df %>% 
      mutate(category = cat)
  }
) %>%
  rename(journal = X1) %>%
  mutate(journal = str_remove(journal, "^[0-9]*\\. "))
```


Preview:

```{r preview journals_and_cat}
journals_df %>% head()
```

Number of distinct journals:

```{r n_unique_journals}
n_journals <- distinct(journals_df %>% select(journal)) %>% nrow
n_journals
```

# Overview and basic statistics

## Category statistics

Number of journals in each category:

```{r n_journals_cat}
journals_df %>%
  group_by(category) %>%
  count() %>%
  arrange(desc(n))
```

Basic statistics (minimum, maximum, first quartile, median, mean, third quartile):

```{r average_st_dev_cat}
journals_df %>%
  group_by(category) %>%
  count() %>%
  summary()
```

## Journals statistics

A journal can be linked to more categories: for this reason the number of journals is much lower than number of rows of `journals_df`. We calculate for each journal how many categories it belongs to:

```{r journals_in_categories}
n_cat_per_journals <- 
  journals_df %>%
  group_by(journal) %>%
  count() %>%
  arrange(desc(n))
```

Preview:

```{r n_cat_per_journals}
n_cat_per_journals %>% head(n = 10)
```

Number of journals in two or more categories:

```{r multidisciplinarity}
n_cat_per_journals %>%
  filter(n > 1) %>%
  nrow()
```

in percentage:

```{r}
n_cat_per_journals %>%
  filter(n > 1) %>%
  nrow() / n_journals * 100
```

Maximum number of categories a journal belongs to:

```{r}
max_categories <- max(n_cat_per_journals$n)
max_categories
```

We spread the name of the journals to columns:

```{r expand_columns}
journals_df_expand <- journals_df %>%
  group_by(journal) %>%
  summarize(categories = paste(category, collapse = ";")) %>%
  separate(categories, 
           into = map_chr(
             seq(1, max_categories), function(n) {
               str_c("category", n, collapse = "_")
               }
             ),
           sep = ";",
           extra = "merge",
           fill = "right"
)
journals_df_expand
```

# Save output

## Save tidy dataframe

We save the tidy dataframe `journals_df` as tab separated text file, `journal_category_tidy.tsv`, in `data/processed`:

```{r tidy_df_save}
journals_df %>%
  write_tsv(
    path = here("data", "processed", "journal_category_tidy.tsv"),
    na = "")
```

It will be the start point of any further data analysis.

## Save expanded dataframe

We save also the expanded dataframe in tab separated text file `category_of_journals_expanded.tsv` in folder `data/processed`:

```{r expand_df}
journals_df_expand %>%
  write_tsv(
    path = here("data", "processed", "category_of_journals_expanded.tsv"),
    na = "")
```

This file is a better overview.
